Anomaly detection with kernel density estimation on manifolds

### 2022-03-21 paper/paper.Rmd
Definitions for differentiable manifolds and tangent space

# Variable kernel density estimation with manifold embeddings {#vkde}

@@

## Algorithm for outlier detection on manifold learning embedding

@@

# Simulation {#simulation}

>>>>

# Notions about Riemannian geometry {#riemgeo}

In this section, we present some notions about the Riemannian geometry used in this paper.

## Differentiable manifolds

In topology, a *homeomorphism* is a bijective map between two topological spaces that is continuous in both directions. 
A *Hausdorff space* is a topological space where any two distinct points can be separated by disjoint neighborhoods.
A $d$-dimensional (topological) *manifold* $M$ is a connected Hausdorff space $(M, \mathcal{T}_M)$ where the neighborhood $U$ for each point $p$ is homeomorphic to an open subset $V$ of the Euclidean space $\mathbb{R}^d$. Such a homeomorphism $\varphi: U \rightarrow V$ and $U$ gives a (coordinate) *chart*, denoted as $(U, \varphi)$, with the corresponding local coordinates $(x_1(p),\dots, x_d(p)) := \varphi(p)$. A collection of charts $\{U_\alpha, \varphi_\alpha\}$ ranging over the manifold $M$ is called an *atlas*, denoted as $\mathcal{A}$.
The manifold $M$ is a *differentiable manifold* if there exists an atlas of $M$, $\{U_\alpha, \varphi_\alpha\}$, such that the *transition maps* between any two charts,
$$
\varphi_\beta \circ \varphi_\alpha^{-1}: \varphi_\alpha(U_\alpha \cap U_\beta) \rightarrow \varphi_\beta(U_\alpha \cap U_\beta),
$$
are differentiable of class $C^\infty$ (smooth).

## Tangent vector and tangent space

The tangent vector at point $p$ can be intuitively viewed as the velocity of a curve passing through the point $p$ or as the directional derivatives at $p$. Here we define the tangent vector via the velocity of curves.

For any $p \in M$, let $\gamma_1: (-\epsilon_1, \epsilon_1)\rightarrow M$ and $\gamma_2: (-\epsilon_2, \epsilon_2)\rightarrow M$ be two smooth curve through $p$, i.e. $\gamma_1(0) = \gamma_2(0) = p$. $\gamma_1$ and $\gamma_2$ are equivalent if and only if there exists a chart, $(U,\varphi)$, at $p$ such that
$$
(\varphi \circ \gamma_1)^\prime(0) = (\varphi \circ \gamma_2)^\prime(0).
$$
A *tangent vector* to $M$ at $p$ is any equivalent class of the differentiable curves initialized at $p$. 
The set of all tangent vectors at $p$ defines the *tangent space* of $M$ at $p$, denoted as $T_pM$. The tangent space is vector space of dimension $d$, equal to the dimension of $M$, and it does not depend on the charts $\varphi$ locally at $p$. The collection of all tangent spaces defines the *tangent bundle* $TM = \cup_{p \in M}T_pM$.
Tangent vectors can also be seen as the directionnal derivatives at $p$. For a given coordinate chart $\varphi=(x_1,\dots,x_d)$, the tangent vectors defining partial derivatives are denoted as $\frac{\partial}{\partial x_1}(p),\dots,\frac{\partial}{\partial x_d}(p)$, which define a *basis* of the tangent space $T_pM$.

The tangent space $T_pM$ also admits a dual space $T^*_pM$ called the *cotangent space* with corresponding *cotangent vector* $z_p: T^*_pM \rightarrow \mathbb{R}^d$, and basis denoted as $dx_1(p),\dots,dx_d(p)$.

## Pushforward and pullback metric

embedding pushforward metric [@Perrault-Joncas2013-pq]

## Riemannian metrics

Let $(\mathcal{M}, g)$ be a $d$-dimensional Riemannian manifold without boundary. Denote $d_g$ the Riemannian distance. Assume that $(\mathcal{M}, g)$ is complete, i.e. $(\mathcal{M}, d_g)$ is a complete metric space, and that the injectivity radius of $(\mathcal{M}, g)$, $inj_g\mathcal{M}$, is strictly positive.

Denote $B_s(p)$ the normal ball in $(\mathcal{M}, g)$ centered at point $p$ with raidus $s$.

$d_g$ is induced by $g$ and by the injectivity radius of $(\mathcal{M}, g)$.

normal ball

A geodesic is a smooth curve that locally joins their points along the shortest path. All the geodesics passing through p are called radial geodesics.

## Exponential map and logarithmic map

## Canonical metric

For a compact Kahler manifold M, canonical metric can mean: (i) Kahler-Enstein metric or (ii) metric with constant scalar curvature which is the extremal metric of the Calabi functional.

## Isometry

## Laplace-Beltrami operator

## Volumn density function

# Kernel density estimation on manifolds {#vkde}

@@

# Outlier detection on manifold learning embedding

@@

# Simulations {#simulation}

### 2022-03-21 paper/paper.tex
Definitions for differentiable manifolds and tangent space

\hypertarget{vkde}{%
\section{Variable kernel density estimation with manifold embeddings}\label{vkde}}

@@

\hypertarget{algorithm-for-outlier-detection-on-manifold-learning-embedding}{%
\subsection{Algorithm for outlier detection on manifold learning embedding}\label{algorithm-for-outlier-detection-on-manifold-learning-embedding}}

@@

\hypertarget{simulation}{%
\section{Simulation}\label{simulation}}

>>>>

\hypertarget{riemgeo}{%
\section{Notions about Riemannian geometry}\label{riemgeo}}

In this section, we present some notions about the Riemannian geometry used in this paper.

A geodesic is a smooth
curve that locally joins their points along the shortest path. All
the geodesics passing through p are called radial geodesics. The
local coordinates defined by the chart ðU ; exp1 Þ are called p
(Riemannian) normal coordinates with center p.~Note that normal coordinates preserve the distances on radial geode- sics. For example, unfolding a sphere onto a plane in normal coordinates can preserve the distances on great circles, like paring an orange. Note that for a manifold with complex topology, for example, a torus, multiple normal coordinate charts, rather than one single chart, are needed to cover the whole manifold.

\hypertarget{differentiable-manifolds}{%
\subsection{Differentiable manifolds}\label{differentiable-manifolds}}

In topology, a \emph{homeomorphism} is a bijective map between two topological spaces that is continuous in both directions.
A \emph{Hausdorff space} is a topological space where any two distinct points can be separated by disjoint neighborhoods.
A \(d\)-dimensional (topological) \emph{manifold} \(M\) is a connected Hausdorff space \((M, \mathcal{T}_M)\) where the neighborhood \(U\) for each point \(p\) is homeomorphic to an open subset \(V\) of the Euclidean space \(\mathbb{R}^d\). Such a homeomorphism \(\varphi: U \rightarrow V\) and \(U\) gives a (coordinate) \emph{chart}, denoted as \((U, \varphi)\), with the corresponding local coordinates \((x_1(p),\dots, x_d(p)) := \varphi(p)\). A collection of charts \(\{U_\alpha, \varphi_\alpha\}\) ranging over the manifold \(M\) is called an \emph{atlas}, denoted as \(\mathcal{A}\).
The manifold \(M\) is a \emph{differentiable manifold} if there exists an atlas of \(M\), \(\{U_\alpha, \varphi_\alpha\}\), such that the \emph{transition maps} between any two charts,
\[
\varphi_\beta \circ \varphi_\alpha^{-1}: \varphi_\alpha(U_\alpha \cap U_\beta) \rightarrow \varphi_\beta(U_\alpha \cap U_\beta),
\]
are differentiable of class \(C^\infty\) (smooth).

\hypertarget{tangent-vector-and-tangent-space}{%
\subsection{Tangent vector and tangent space}\label{tangent-vector-and-tangent-space}}

The tangent vector at point \(p\) can be intuitively viewed as the velocity of a curve passing through the point \(p\) or as the directional derivatives at \(p\). Here we define the tangent vector via the velocity of curves.

For any \(p \in M\), let \(\gamma_1: (-\epsilon_1, \epsilon_1)\rightarrow M\) and \(\gamma_2: (-\epsilon_2, \epsilon_2)\rightarrow M\) be two smooth curve through \(p\), i.e.~\(\gamma_1(0) = \gamma_2(0) = p\). \(\gamma_1\) and \(\gamma_2\) are equivalent if and only if there exists a chart, \((U,\varphi)\), at \(p\) such that
\[
(\varphi \circ \gamma_1)^\prime(0) = (\varphi \circ \gamma_2)^\prime(0).
\]
A \emph{tangent vector} to \(M\) at \(p\) is any equivalent class of the differentiable curves initialized at \(p\).
The set of all tangent vectors at \(p\) defines the \emph{tangent space} of \(M\) at \(p\), denoted as \(T_pM\). The tangent space is vector space of dimension \(d\), equal to the dimension of \(M\), and it does not depend on the charts \(\varphi\) locally at \(p\). The collection of all tangent spaces defines the \emph{tangent bundle} \(TM = \cup_{p \in M}T_pM\).
Tangent vectors can also be seen as the directionnal derivatives at \(p\). For a given coordinate chart \(\varphi=(x_1,\dots,x_d)\), the tangent vectors defining partial derivatives are denoted as \(\frac{\partial}{\partial x_1}(p),\dots,\frac{\partial}{\partial x_d}(p)\), which define a basis of the tangent space \(T_pM\).

The tangent space \(T_pM\) also admits a dual space \(T^*_pM\) called the cotangent space with corresponding cotangent vector \(z_p: T^*_pM \rightarrow \mathbb{R}^d\), and basis denoted as \(dx_1(p),\dots,dx_d(p)\).

Define the directional derivative \(D_v\) in the direction of \(v\) as
\[
\forall f \in C^\infty(M): D_v(f)
\]

\hypertarget{riemannian-metrics}{%
\subsection{Riemannian metrics}\label{riemannian-metrics}}

Let \((\mathcal{M}, g)\) be a \(d\)-dimensional Riemannian manifold without boundary. Denote \(d_g\) the Riemannian distance. Assume that \((\mathcal{M}, g)\) is complete, i.e.~\((\mathcal{M}, d_g)\) is a complete metric space, and that the injectivity radius of \((\mathcal{M}, g)\), \(inj_g\mathcal{M}\), is strictly positive.

Denote \(B_s(p)\) the normal ball in \((\mathcal{M}, g)\) centered at point \(p\) with raidus \(s\).

\(d_g\) is induced by \(g\) and by the injectivity radius of \((\mathcal{M}, g)\).

normal ball

\hypertarget{exponential-map-and-logarithmic-map}{%
\subsection{Exponential map and logarithmic map}\label{exponential-map-and-logarithmic-map}}

\hypertarget{pushforward-and-pullback-metric}{%
\subsection{Pushforward and pullback metric}\label{pushforward-and-pullback-metric}}

embedding pushforward metric \autocite{Perrault-Joncas2013-pq}

\hypertarget{canonical-metric}{%
\subsection{Canonical metric}\label{canonical-metric}}

For a compact Kahler manifold M, canonical metric can mean: (i) Kahler-Enstein metric or (ii) metric with constant scalar curvature which is the extremal metric of the Calabi functional.

\hypertarget{isometry}{%
\subsection{Isometry}\label{isometry}}

\hypertarget{laplace-beltrami-operator}{%
\subsection{Laplace-Beltrami operator}\label{laplace-beltrami-operator}}

\hypertarget{volumn-density-function}{%
\subsection{Volumn density function}\label{volumn-density-function}}

\hypertarget{vkde}{%
\section{Kernel density estimation on manifolds}\label{vkde}}

@@

\hypertarget{outlier-detection-on-manifold-learning-embedding}{%
\section{Outlier detection on manifold learning embedding}\label{outlier-detection-on-manifold-learning-embedding}}

@@

\hypertarget{simulation}{%
\section{Simulations}\label{simulation}}
